Data analysis

### 2025-03-05 processing/01-proc-data.qmd
Update analysis with income and gaps

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(perc_sal_gerente, perc_sal_obrero),
           .fns = ~ sjlabelled::set_na(., na = 0)
      )
    )

@@

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(just_sal_gerente, just_sal_obrero),
           .fns = ~ sjlabelled::set_na(., na = 0)
      )
    )

@@

db_proc$perc_inequality <- sjlabelled::set_label(db_proc$perc_inequality, 
                        label = "Inequality gap justification")

@@

>>>>

db_proc <- db_proc %>% 
  mutate(perc_sal_obrero=replace(perc_sal_obrero, perc_sal_obrero <= 40000  | perc_sal_obrero>= 1000000, NA)) %>%
  mutate(perc_sal_gerente=replace(perc_sal_gerente, perc_sal_gerente <= 250000 | perc_sal_gerente>= 100000001, NA))

@@

db_proc <- db_proc %>% 
  mutate(just_sal_obrero=replace(just_sal_obrero, just_sal_obrero <= 40000  | just_sal_obrero>=999999999999999, NA)) %>%
  mutate(just_sal_gerente=replace(just_sal_gerente, just_sal_gerente <= 100000 | just_sal_gerente>=999999999999999, NA))

@@

db_proc$just_inequality <- sjlabelled::set_label(db_proc$just_inequality, 
                        label = "Inequality gap justification")

@@

```{r}
# Household income----------------------------------------------------------------

# Ltop is the lower limit of the top category, Ltop-1 is the lower limit of the category before the
# top one, ftop is the frequency in the top category, and ftop-1 is the frequency in the
# category before the top one.

sjmisc::frq( elsoc_long_2016_2023$m30)
Ltop_1<- 1850001
Ltop  <- 2700000

ftop_1 <- 42
ftop   <- 34

V  = (log(ftop_1 + ftop) - log(ftop)) / (log(Ltop) -log(Ltop_1))
M_top = 0.5* Ltop *(1+(V/(V-1)));M_top # = $3.897.232

#Impute midpoint of income ranges
elsoc_long_2016_2023$m30_rec <-
  as.numeric(car::recode( elsoc_long_2016_2023$m30,
                         "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=3897232;NA=NA;c(-888,-999)=NA"))

summary( elsoc_long_2016_2023$m30_rec)

#Impute midpoint of income ranges (2021)

frq(elsoc_long_2016_2023$m30b)
Ltop_1<- 450000
Ltop  <- 700000
ftop_1 <- 18
ftop   <- 27

V  = (log(ftop_1 + ftop) - log(ftop)) / (log(Ltop) -log(Ltop_1))
M_top = 0.5* Ltop *(1+(V/(V-1))) # = $2.941.412

 elsoc_long_2016_2023$m30b_rec <-
  as.numeric(car::recode( elsoc_long_2016_2023$m30b,
                         "1=125000;2=300000;3=400000;4=575000;5=2941412;NA=NA;c(-888,-999)=NA"))
summary( elsoc_long_2016_2023$m30b_rec)

# Combine m30_rec with m30b_rec
 elsoc_long_2016_2023$m30_rec <- 
  ifelse(is.na( elsoc_long_2016_2023$m30_rec),
         yes =  elsoc_long_2016_2023$m30b_rec,
         no =  elsoc_long_2016_2023$m30_rec)
summary( elsoc_long_2016_2023$m30_rec)

#Recode DK/DA of Income to NA
elsoc_long_2016_2023$m29_rec <-
  as.numeric(car::recode( elsoc_long_2016_2023$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_long_2016_2023$m29_imp <- 
  ifelse(test = !is.na( elsoc_long_2016_2023$m29_rec),
         yes =   elsoc_long_2016_2023$m29_rec,
         no =   elsoc_long_2016_2023$m30_rec)
summary( elsoc_long_2016_2023$m29_imp)

#  elsoc_long_2016_2023$m29_imp <- 
#   ifelse(test = is.na( elsoc_long_2016_2023$m29_imp),
#          yes =   elsoc_long_2016_2023$m30b_rec,
#          no =   elsoc_long_2016_2023$m29_imp)

summary( elsoc_long_2016_2023$m29_imp)

# N Household:
elsoc_long_2016_2023 <-
   elsoc_long_2016_2023 %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54,
                            ola == 6 ~ m54,
                            ola == 7 ~ m54))

summary( elsoc_long_2016_2023$n_hogar)
#imputar tamanio del hogar de ola 5 a ola 6
elsoc_long_2016_2023 <- 
   elsoc_long_2016_2023 %>%  
  group_by(idencuesta) %>%
  mutate(n_hogar = if_else(ola == 6 & is.na(n_hogar), last(n_hogar[ola == 5]), n_hogar)) %>% 
  ungroup()

sjmisc::frq( elsoc_long_2016_2023$n_hogar)
table( elsoc_long_2016_2023$n_hogar, elsoc_long_2016_2023$ola)

#Recode DK/DA to NA
 elsoc_long_2016_2023$n_hogar_r<-
  car::recode( elsoc_long_2016_2023$n_hogar,"c(-888,-999)=NA")

table( elsoc_long_2016_2023$m13, elsoc_long_2016_2023$n_hogar_r)

#ingresos individuales para hogares de 1 persona
 elsoc_long_2016_2023 %>% sjmisc::find_var('Ingreso')
 elsoc_long_2016_2023$inc_ind <- ifelse( elsoc_long_2016_2023$n_hogar_r==1, elsoc_long_2016_2023$m13,NA)
summary( elsoc_long_2016_2023$inc_ind)
# 
# c(40000,
# (40001 + 85000) / 2,
# (85001 + 125000) / 2,
# (125001 + 170000) / 2,
# (170001 + 210000) / 2,
# (210001 + 230000) / 2,
# (230001 + 280000) / 2,
# (280001 + 320000) / 2,
# (320001 + 360000) / 2,
# (360001 + 400000) / 2,
# (400001 + 465000) / 2,
# (465001 + 540000) / 2,
# (540001 + 665000) / 2,
# (665001 + 850000) / 2,
# (850001 + 1300000) / 2,
# 1300001)

frq( elsoc_long_2016_2023$m14)
Ltop_1<- 850001
Ltop  <- 1300001 
ftop_1 <- 65
ftop   <- 40

V  = (log(ftop_1 + ftop) - log(ftop)) / (log(Ltop) -log(Ltop_1))
M_top = 0.5* Ltop *(1+(V/(V-1))) # = $1.811.247

 elsoc_long_2016_2023$m14_rec<- 
  car::recode( elsoc_long_2016_2023$m14,
              "1=40000.0;2=62500.5;3=105000.5;4=147500.5;5=190000.5;6=220000.5;
              7=255000.5;8=300000.5;9=340000.5;10=380000.5;11=432500.5;12=502500.5;
              13=602500.5;14=757500.5;15=1075000.5;16=1811247"
  )

 elsoc_long_2016_2023$inc_ind <- ifelse( elsoc_long_2016_2023$n_hogar_r==1 & is.na( elsoc_long_2016_2023$inc_ind), elsoc_long_2016_2023$m14_rec, elsoc_long_2016_2023$inc_ind)
summary( elsoc_long_2016_2023$inc_ind)

# Ingreso del hogar pegarle ingreso del hogar unipersonal
summary( elsoc_long_2016_2023$m29_imp)
 elsoc_long_2016_2023$m29_imp <- 
  ifelse(is.na( elsoc_long_2016_2023$m29_imp) &  elsoc_long_2016_2023$n_hogar_r==1,
          elsoc_long_2016_2023$inc_ind,
          elsoc_long_2016_2023$m29_imp) #recupero 14 casos
summary( elsoc_long_2016_2023$m29_imp)

# Household Equivalent income 
# 1.0 to the first adult;
# 0.5 to the second and each subsequent person aged 14 and over;
# 0.3 to each child aged under 14.

 elsoc_long_2016_2023$n_hogar_r_eq <- ( elsoc_long_2016_2023$n_hogar_r + 1) /2
table( elsoc_long_2016_2023$n_hogar_r_eq)

frq( elsoc_long_2016_2023$n_hogar_r_eq)

# Per capita household income:
 elsoc_long_2016_2023$ing_pc <- 
  (elsoc_long_2016_2023$m29_imp/ elsoc_long_2016_2023$n_hogar_r)

# Per capita household income equivalized (OECD):
 elsoc_long_2016_2023$ing_pc_eq <- 
  ( elsoc_long_2016_2023$m29_imp/ elsoc_long_2016_2023$n_hogar_r_eq)

descr( elsoc_long_2016_2023$ing_pc_eq)
descr( elsoc_long_2016_2023$ing_pc)
 elsoc_long_2016_2023$ing_pc <-
  sjlabelled::set_label(x =  elsoc_long_2016_2023$ing_pc,
                        label = "Household income per capita")

sjmisc::descr( elsoc_long_2016_2023$ing_pc)

# Compute income quintiles
 elsoc_long_2016_2023 <-  elsoc_long_2016_2023 %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

 elsoc_long_2016_2023$quintil <- 
  factor( elsoc_long_2016_2023$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

elsoc_long_2016_2023$quintil <- 
  sjlabelled::set_label(x =  elsoc_long_2016_2023$quintil,
                        label = "Household income quintile per capita")

sjmisc::frq( elsoc_long_2016_2023$quintil)

#include new quintile category with missing cases
 elsoc_long_2016_2023$quintil1<-
  car::recode( elsoc_long_2016_2023$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

# elsoc_long_2016_2023$quintil1 <- factor( elsoc_long_2016_2023$quintil1, c("Q1","Q2","Q3","Q4","Q5","QNA"))

 elsoc_long_2016_2023$quintil1 <- 
  sjlabelled::set_label(x =  elsoc_long_2016_2023$quintil1,
                        label = "Household income quintile per capita (NA)") 
sjmisc::frq( elsoc_long_2016_2023$quintil1)

```{r}
elsoc_income <- elsoc_long_2016_2023 %>% 
  dplyr::select(idencuesta, ola, ing_pc, quintil, quintil1) %>% 
  dplyr::filter(ola %in% c(1,2,3,4,6,7))

db_proc <- left_join(db_proc, elsoc_income, by = c("idencuesta", "wave" = "ola"))

### 2025-03-05 processing/02-analysis.qmd
Update analysis with income and gaps

m7 <- lmer(mjp ~ 1 + perc_inequality + just_inequality + 
             merit_effort + merit_talent + perc_inequality_mean +
             just_inequality_mean + merit_effort_mean + 
             merit_talent_mean + age + sex + ess + ideo + educ + wave_f + (1 | idencuesta),
                data = df1)
```

@@

ccoef <- list(
  "(Intercept)" = "Intercept",
  perc_inequality = "Perception inequality (WE)",
  just_inequality = "Justification inequality (WE)",
  merit_effort = "Merit: Effort (WE)",
  merit_talent = "Merit: Talent (WE)",
  perc_inequality_mean = "Perception inequality (BE)",
  just_inequality_mean = "Justification inequality (BE)",
  merit_effort_mean = "Merit: Effort (BE)",
  merit_talent_mean = "Merit: Talent (BE)",
  "age30-49" = "30-49",
  "age50-64" = "50-64",
  "age65 or more" = "65 or more",
  sexFemale = "Female (Ref.= Male)",
  ess = "Subjective social status",
  ideoCenter = "Center",
  ideoRight = "Right",
  "ideoDoes not identify" = "Does not identify",
  educUniversitary = "Universitary (Ref.= Less than universitary)",
  wave_f2017 = "2017",
  wave_f2018 = "2018",
  wave_f2019 = "2019",
  wave_f2022 = "2022",
  wave_f2023 = "2023"
  )

texreg::htmlreg(list(m1, m2, m3, m4, m5, m6, m7),
               custom.model.names = c(paste0("Model ", seq(1:7))),
               caption = NULL,
               stars = c(0.05, 0.01, 0.001),
               custom.coef.map = ccoef,
               digits = 3,
               groups = list("Age (Ref. = 18-29)" = 10:12,"Political identification (Ref.= Left)" = 15:17, "Wave (Ref.= 2016)" = 19:23),
               custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars. \\\\ \\item Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 4364).",
               threeparttable = T,
               leading.zero = T,
               float.pos = "h!",
               use.packages = F,
               booktabs = T,
               scalebox = 1)
```

>>>>

m7 <- lmer(mjp ~ 1 + perc_inequality + just_inequality + 
             merit_effort + merit_talent + perc_inequality_mean +
             just_inequality_mean + merit_effort_mean + 
             merit_talent_mean + ideo + quintil + educ + ess + age + sex + wave_f + (1 | idencuesta),
                data = df1)
```

@@

ccoef <- list(
  "(Intercept)" = "Intercept",
  perc_inequality = "Perception inequality (WE)",
  just_inequality = "Justification inequality (WE)",
  merit_effort = "Merit: Effort (WE)",
  merit_talent = "Merit: Talent (WE)",
  perc_inequality_mean = "Perception inequality (BE)",
  just_inequality_mean = "Justification inequality (BE)",
  merit_effort_mean = "Merit: Effort (BE)",
  merit_talent_mean = "Merit: Talent (BE)",
  ideoCenter = "Center",
  ideoRight = "Right",
  "ideoDoes not identify" = "Does not identify",
  quintilQ2 = "Quintile 2",
  quintilQ3 = "Quintile 3",
  quintilQ4 = "Quintile 4",
  quintilQ5 = "Quintile 5",
  educUniversitary = "Universitary (Ref.= Less than universitary)",
  ess = "Subjective social status",
  "age30-49" = "30-49",
  "age50-64" = "50-64",
  "age65 or more" = "65 or more",
  sexFemale = "Female (Ref.= Male)",
  wave_f2017 = "2017",
  wave_f2018 = "2018",
  wave_f2019 = "2019",
  wave_f2022 = "2022",
  wave_f2023 = "2023"
  )

texreg::htmlreg(list(m1, m2, m3, m4, m5, m6, m7),
               custom.model.names = c(paste0("Model ", seq(1:7))),
               caption = NULL,
               stars = c(0.05, 0.01, 0.001),
               custom.coef.map = ccoef,
               digits = 3,
               groups = list("Political identification (Ref.= Left)" = 10:12, "Income (Ref.= Quintile 1)" = 13:16, "Age (Ref. = 18-29)" = 19:21, "Wave (Ref.= 2016)" = 23:26),
               custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars. \\\\ \\item Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 4257).",
               threeparttable = T,
               leading.zero = T,
               float.pos = "h!",
               use.packages = F,
               booktabs = T,
               scalebox = 1)
```
